Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-04-23 article.tex
Simplify Makefile
Describe in silico TMH AASM determination

@@

\richel{Check against stupidity}
The human immune system and human pathogen are in an evolutionary
arms race, in which our immune systems is selected for the detection
of pathogens, whereas pathogens are selected to avoid being detected.
If the human immune system is more prone to display (and possibly 
detect) polypeptides derived from TMHs, one may expect to find 
an evolutionary signal in pathogens as well.

@@

\richel{Check against stupidity}
Regarding human viruses, we do not know the selection pressure
exerted by the human immune system.
We do know that this selection pressure is 
limited for seasonal influenza \cite{han2019individual},
but this study does not show how frequent selected upon 
TMHs \richel{Use this study, if COVID-19 dataset is too weak}.
For COVID-19, there is preliminary evidence that the strongest
selection pressure is upon ORF3 (an ion channel and a viroporin)
and ORF8 (unknown function). ORF3 has 3 TMHs that appear irrelevant for
the natural selection of COVID-19.
We predict that evolutionary conservation is strongest 
for regions that are essential for viral proteins'
functions (such as catalytic binding sites), instead of TMHs.
Because these essential regions are rare in a viral proteome,
and differ widely between pathogens, we expect that the human immune system
is not tailored to detect these sites, similar to the aforementioned
influenza study. 
TMHs, however, have less 
functional constraints compared to catalytic regions,
and as such are more general. 
Because TMHs are more general,
we expect it for the immune system to pay off to be more attentive towards
TMHs. 
Additionally, viruses are not only selected for their evasion of an 
immune response, yet are selected to have a high reproductive value.
Because the strength of the evolutionary signal is unknown,
we use a strong data set to be able to detect it.

@@

\richel{Check against stupidity}
Regarding human pathogenic bacteria, we do not know the extent to which the 
human immune systems selects on these, nor which loci are selected
upon \richel{so I should read the literature on that}.
Because most bacterial pathogens are generalists, that is,
can infect multiple hosts, we expect the selection pressure exerted
by the human immune response to be weak.
Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
Due to this, the strength of the evolutionary signal is unknown,
thus we need a strong data set to be able to detect it.

@@

>>>>

% Just try to parse, do not ask for input
\nonstopmode

@@

\richel{Check against stupidity}
The human immune system and human pathogen are in an evolutionary
arms race, in which our immune systems is selected for the detection
of pathogens, whereas pathogens are selected to avoid being detected.
If the human immune system is more prone to bind hydrophobic polypeptides, 
most of which derived from TMHs, one may expect to find 
an evolutionary signal in TMHs.

\paragraph{Amino acid substitution models}

One important assumption for such phylogenetic inference
is the amino acid substitution model (AASM), which may be different
in TMHs. An AASM encodes the relative rates at which amino acids
change through evolutionary time. Some amino acids have more 
DNA triplets encoding for them, which is reflected in the AASM.
Already for decades, AASMs are devised and refined from whole proteomes,
such as the Blosum62 (\cite{henikoff1992amino}), 
Dayhoff (\cite{dayhoff197822}), JTT (\cite{jones1992rapid}), 
and WAG (\cite{whelan2001general}) models. 
That such a generalization
is not always desired, AASMs are devised for 
chloroplasts (the CPREV model, see \cite{adachi2000plastid})
and mitochondria (the MTREV24 model, see \cite{adachi1996model}).
In all AASM, TMHs are generalized out and therefore, 
there is no specialized AASM for TMHS. It would, however,
be possible to derive an AASM from in silico predictions. 
An in silico prediction would bypass any effects of evolution.
Another AASM could be constructed from humans, if we assume
that there is no selection on TMHs, except for being functional.

@@

\richel{Check against stupidity}
Regarding human viruses, we do not know the selection pressure
exerted by the human immune system.
We do know that this selection pressure is 
limited for seasonal influenza \cite{han2019individual},
but this study does not look into the evolution of TMHs
\richel{Use this study, if COVID-19 dataset is too weak}.
For COVID-19, there is preliminary evidence that the strongest
selection pressure is upon ORF3 (an ion channel and a viroporin)
and ORF8 (unknown function). ORF3 has 3 TMHs that appear irrelevant for
the natural selection of COVID-19.
We predict that evolutionary conservation is strongest 
for regions that are essential for viral proteins'
functions (such as catalytic binding sites), instead of TMHs.
Because these essential regions are rare in a viral proteome,
and differ widely between pathogens, we expect that the human immune system
is not tailored to detect these sites, similar to the aforementioned
influenza study. 
TMHs, however, have less 
functional constraints compared to catalytic regions,
and as such are more general. 
Because TMHs are more general,
we expect it for the immune system to pay off to be more attentive towards
TMHs.
For viral TMHs that are undetected by most human haplotypes, we
expect these to be conserved, to remain avoiding detection.
For viral TMHs that are detected by most human haplotypes, we
expect there to be a higher mutation rate, to avoid detection.
Additionally, viruses are not only selected for their evasion of an 
immune response, yet are selected to have a high reproductive value.
Because the strength of the evolutionary signal is unknown,
we use a strong data set to be able to detect it.

@@

\richel{Check against stupidity}
Regarding human pathogenic bacteria, we do not know the extent to which the 
human immune systems selects on these, nor which loci are selected
upon \richel{so I should read the literature on that}.
Because most bacterial pathogens are generalists, that is,
can infect multiple hosts, we expect the selection pressure exerted
by the human immune response to be weak.
For bacterial TMHs that are undetected by most human haplotypes, we
expect these to be conserved, to remain avoiding detection.
For bacterial TMHs that are detected by most human haplotypes, we
expect there to be a higher mutation rate, to avoid detection.
Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
Due to this, the strength of the evolutionary signal is unknown,
thus we need a strong data set to be able to detect it.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{In silico prediction of TMH AASM}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We use a Monte-Carlo approach to produce an AASM for TMHs.
We simulate a random amino acid sequence with a length 23.
Using TMHMM, we determine if a sequence encodes for a TMH.
For a TMH sequence, we independently change 
each locus to all other amino acids.
Of each mutation, we count how often the new sequence remains a TMH.
Doing this for many random amino acid sequences, we can determine
the AASM. See figure \ref{fig:tmh_aasm_method} for a schematic
overview of this method.

\begin{figure}[!htbp]
  \includegraphics[width=0.9\textwidth]{tmh_aasm_method.png}
  \caption{
    In silico prediction of TMH AASM
  }
  \label{fig:tmh_aasm_method}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Baseline TMH mutation rate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Detecting evolution}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
